---
title: DS-6331 Time Series Project- Covid-19 Death Rate and the Influence of Vaccine
  and other variables
author: "Tai Chowdhury and Simerpreet Reddy"
date: "4/09/2021"
output:
  word_document: default
  html_document: default
---

### Objective:
Analyze  Covid-19 Death Rate(per day) and the influence of other factors- New cases per day, Vaccination Rates, number of hospitalizations and number of ICU patients per day. We will perform the following taks:
1) Performing Univariate analysis on 'New Deaths per million'.
2) Build ARIMA model.
3) Build the following models: We will first build models using only 'New Deaths per million' and 'New Vaccinations per million' and then will analyze all other variables as well. We will also analyse lagged variables. Models to be built: 
  i)Multivariate analysis 
  ii) VAR models
  iii) MLP models
  iv) ensemble models using any of the above 2 models
  
### Data Set: 
1) All Covid: Data Set with all data and related variables since the beginning of the pandemic.
2) Post Vaccine: Daily data since the first vaccine got administered in the US(12/21/20) through 03/26/2021 (96 observations). This is the data set we are bulding the models on. 

### Data Set description:
96 observations with the following attributes: Date, New Cases per million, New Deaths per million, New ICU patients per million, New Hospitalizations per million, New Vaccinations per million

### Source:
Coronavirus (COVID-19) Vaccinations - Statistics and Research - Our World in Data. It is updated daily and includes data on confirmed cases, deaths, hospitalizations, testing, and vaccinations as well as other variables of potential interest.

```{r, message=F, warning=F}
#Include necessary libraries
library(nnfor)
library(forecast)
library(vars)
library(tswge)

All_Covid <- read.csv(file.choose(), header=TRUE)
nrow(All_Covid)

Post_Vaccine <- read.csv(file.choose(), header=TRUE)
nrow(Post_Vaccine)

str(Post_Vaccine)

#Rolling Window Function def.n
Rolling_Window_ASE = function(series, trainingSize, horizon, s, d, phis, thetas)
{
# trainingSize = 70
# horizon = 12
ASEHolder = numeric()
# s = 10
# d = 0
# phis = phis
# thetas = thetas
for( i in 1:(length(series)-(trainingSize + horizon) + 1))
{
  forecasts = fore.aruma.wge(series[i:(i+(trainingSize-1))],phi = phis, theta = thetas, s = s, d = d,n.ahead = horizon)
  ASE = mean((series[(trainingSize+i):(trainingSize+ i + (horizon) - 1)] - forecasts$f)^2)
  ASEHolder[i] = ASE
}
ASEHolder
hist(ASEHolder)
WindowedASE = mean(ASEHolder)
print(horizon)
print(trainingSize)
print("The Summary Statistics for the Rolling Window ASE Are:")
print(summary(ASEHolder))
print(paste("The Rolling Window ASE is: ",WindowedASE))
return(WindowedASE)
}
```


#############
### EDA ####
#############

```{r, message=F, warning=F}
#Plotting time series for all variables since the beginning of the pandemic
plotts.wge(All_Covid$new_cases_per_million)
plotts.wge(All_Covid$icu_patients_per_million)
plotts.wge(All_Covid$hosp_patients_per_million)
plotts.wge(All_Covid$new_vaccinations_smoothed_per_million)
plotts.wge(All_Covid$new_deaths_per_million)

# Plotting time series for all variables Post Vaccine
plotts.wge(Post_Vaccine$new_cases_per_million)
plotts.wge(Post_Vaccine$icu_patients_per_million)
plotts.wge(Post_Vaccine$hosp_patients_per_million)
plotts.wge(Post_Vaccine$new_vaccinations_smoothed_per_million)
plotts.sample.wge(Post_Vaccine$new_deaths_per_million)

```


### ARIMA Model 1 (with seasonal component, s = 7)


```{r, message=F, warning=F}
#ARIMA Models as this data is assumed non-stationary
# Step 1 Check if data is white noise
plotts.sample.wge(Post_Vaccine$new_deaths_per_million)
acf(Post_Vaccine$new_deaths_per_million) 
parzen.wge(Post_Vaccine$new_deaths_per_million)
# This data set is not white noise, Autocorrelation are correlated. Damped sinusoidal indicating seasonality.

#Bulding Models
#ARIMA Model 1 Only Seasonality 

#Take out the s=7 (1-B^7) term from the model
pv_s7 <- artrans.wge(Post_Vaccine$new_deaths_per_million,phi.tr = c(rep(0,6),1)) #Looks statinary

#Plot the timeseries without the seasonal component
plotts.sample.wge(pv_s7)

#Overfitting
est.ar.wge(Post_Vaccine$new_deaths_per_million, p=10, type='burg')

#aic with defaults
aic.wge(pv_s7) #=p=4, q=1
arma_s7_1 <- est.arma.wge(pv_s7, p=4, q=1)

#residuals
plotts.sample.wge(arma_s7_1$res,arlimits = T) #Residuals not white noise

#Check bic
aic5.wge(pv_s7, p=0:13, q=0:3, type="bic")  #BIC picks ARMA(1,0) model. Looking at the data, this does not look appropriate.

#arma_s7_d1 <- est.arma.wge(pv_s7_d1, p=7, q=3)
for_aruma2_s7 = fore.aruma.wge(Post_Vaccine$new_deaths_per_million,s=7,  phi = arma_s7_1$phi,n.ahead = 10, lastn = T)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,96), main = "Last 10 Day forecast with p=7, q=3 and s=7")
lines(seq(87,96,1),for_aruma2_s7$f, col = "red")

#ASE
arma_s7_ase = mean((for_aruma2_s7$f - Post_Vaccine$new_deaths_per_million[(96-10+1):96])^2)
arma_s7_ase

#Rolling_Window_ASE
Rolling_Window_ASE(Post_Vaccine$new_deaths_per_million, trainingSize = 70, horizon = 10, d = 0, phis = arma_s7_1$phi,
                   s= 7, thetas = arma_s7_1$theta)

#Forecasting
#Next 10
for_aruma2_s7_f10 = fore.aruma.wge(Post_Vaccine$new_deaths_per_million,s=7, phi = arma_s7_1$phi,n.ahead = 10)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,101), main = "Next 10 Day forecast with p=7, q=3 and s=7")
lines(seq(97,106,1),for_aruma2_s7_f10$f, col = "red")

#Next 40
for_aruma2_s7_f40 = fore.aruma.wge(Post_Vaccine$new_deaths_per_million,s=7, phi = arma_s7_1$phi,n.ahead = 40)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,136), main = "Next 40 Day forecast with with p=7, q=3 and s=7")
lines(seq(96,135,1),for_aruma2_s7_f40$f, col = "red")

#Next 180
for_aruma2_s7_f180 = fore.aruma.wge(Post_Vaccine$new_deaths_per_million,s=7, phi = arma_s7_1$phi,n.ahead = 180)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,276), main = "Next 180 Day forecast with p=7, q=3 and s=7")
lines(seq(97,276,1),for_aruma2_s7_f180$f, col = "red")

#Next 380
for_aruma2_s7_f380 = fore.aruma.wge(Post_Vaccine$new_deaths_per_million,s=7, phi = arma_s7_1$phi,n.ahead = 380)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,476), main = "Next 10 Day forecast with p=7, q=3 and s=7")
lines(seq(97,476,1),for_aruma2_s7_f380$f, col = "red")
```

### ARMA MODEL 2 - (with (1-B) component)  ####


```{r, message=F, warning=F}
# overfitting to check non stationary portion of the data
est.ar.wge(Post_Vaccine$new_deaths_per_million, p=8, type='burg') #has one 1-0.9868B
# 1st difference
pv_dif1 <- artrans.wge(Post_Vaccine$new_deaths_per_million, phi.tr=1)#dif data does not look like white noise.
plotts.sample.wge(pv_dif1, arlimits = T)
dif1_aic <- aic.wge(pv_dif1) #p=5, q=2
arima_1_diff1 <- est.arma.wge(pv_dif1, p=5, q=2)
#Examine model residuals
plotts.sample.wge(arima_1_diff1$res, arlimits = T) #Not white noise, model needs improvement
#forecast using diff-1
#Last 10
for_aruma1_diff1 = fore.aruma.wge(Post_Vaccine$new_deaths_per_million,d=1, phi = arima_1_diff1$phi,n.ahead = 10, lastn = T)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,96), main = "Last 10 Day forecast with p=5, q=2, s=0 and d=1")
lines(seq(87,96,1),for_aruma1_diff1$f, col = "red")

#Forecast future 10
for_aruma1_diff_f10 = fore.aruma.wge(Post_Vaccine$new_deaths_per_million,d=1, phi = arima_1_diff1$phi,n.ahead = 10)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,106), main = "Next 10 Day forecast with p=5, q=2, s=0 and d=1")
lines(seq(97,106,1),for_aruma1_diff_f10$f, col = "red")

#Forecast future 40
for_aruma1_diff_f40 = fore.aruma.wge(Post_Vaccine$new_deaths_per_million,d=1, phi = arima_1_diff1$phi,n.ahead = 40)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,136), main = "Next 40 Day forecast with p=5, q=2, s=0 and d=1")
lines(seq(96,135,1),for_aruma1_diff_f40$f, col = "red")

#Forecast future 180
for_aruma1_diff_f180 = fore.aruma.wge(Post_Vaccine$new_deaths_per_million,d=1, phi = arima_1_diff1$phi,n.ahead = 180)

#Forecast future 380
for_aruma1_diff_f380 = fore.aruma.wge(Post_Vaccine$new_deaths_per_million,d=1, phi = arima_1_diff1$phi,n.ahead = 380)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,476), main = "Next 380 Day forecast with p=5, q=2, s=0 and d=1")
lines(seq(97,476,1),for_aruma1_diff_f380$f, col = "red")
#ASE
arima1_diff1_ase = mean((for_aruma1_diff1$f - Post_Vaccine$new_deaths_per_million[(96-10+1):96])^2)
arima1_diff1_ase 
Rolling_Window_ASE(Post_Vaccine$new_deaths_per_million, trainingSize = 70, horizon = 10, d = 1, phis = arima_1_diff1$phi,
                   s= 0, thetas = arima_1_diff1$theta) 
```


### ARIMA - (with both seasonal(s = 7) and trend(1-B) components)  ####


```{r, message=F, warning=F}
#Seasonality plus 1-B
plotts.sample.wge(pv_dif1)
pv_s7_d1 <- artrans.wge(pv_dif1,phi.tr = c(rep(0,6),1)) #Looks statinary
plotts.sample.wge(pv_s7_d1)
#Overfitting
est.ar.wge(Post_Vaccine$new_deaths_per_million, p=8, type='burg')
#aic with defaults
aic.wge(pv_s7_d1) #=p=0 q=2
#Check bic
aic5.wge(pv_s7_d1, p=0:13, q=0:3, type="bic")  #BIC picks MA(2) model. Looking at the data, this does not look appropriate.
#aic with bigger p and q range
aic.wge(pv_s7_d1, p=0:13, q=0:3) #=p=7 q=0
arima_s7_d1 <- est.ar.wge(pv_s7_d1, p=7, type='burg')
#residuals
plotts.sample.wge(arima_s7_d1$res, arlimits = T) #Much better model residuals

#Forecast last 10
for_aruma2_s7d1 = fore.aruma.wge(Post_Vaccine$new_deaths_per_million,d=1,s=7,  phi = arima_s7_d1$phi,n.ahead = 10, lastn = T)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,96), main = "Last 10 Day forecast with p=7, q=0, s=7 and d=1")
lines(seq(87,96,1),for_aruma2_s7d1$f, col = "red")

#Forecast last 16
for_aruma2_s7d1_16 = fore.aruma.wge(Post_Vaccine$new_deaths_per_million,d=1,s=7,  phi = arima_s7_d1$phi,n.ahead = 16, lastn = T)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,96), main = "Last 16 Day forecast with p=7, q=0, s=7 and d=1")
lines(seq(81,96,1),for_aruma2_s7d1_16$f, col = "red")
#ASE
arma_s7_d1_ase = mean((for_aruma2_s7d1_16$f - Post_Vaccine$new_deaths_per_million[(96-16+1):96])^2)
arma_s7_d1_ase 
#Rolling_Window_ASE
Rolling_Window_ASE(Post_Vaccine$new_deaths_per_million, trainingSize = 70, horizon = 10, d = 1, phis = arima_s7_d1$phi,
                   s= 7, theta = 0)

#Future forecasts
for_aruma2_s7d1_f14 = fore.aruma.wge(Post_Vaccine$new_deaths_per_million,d=1, s=7, phi = arima_s7_d1$phi,n.ahead = 14)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,110), main = "Next 14 Day forecast with p=7, q=0, s=7 and d=1")
lines(seq(97,110,1),for_aruma2_s7d1_f14$f, col = "red")
for_aruma2_s7d1_f40 = fore.aruma.wge(Post_Vaccine$new_deaths_per_million,d=1,s=7, phi = arima_s7_d1$phi,n.ahead = 40)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,136), main = "Next 40 Day forecast with p=7, q=0, s=7 and d=1")
lines(seq(96,135,1),for_aruma2_s7d1_f40$f, col = "red")
for_aruma2_s7d1_f180 = fore.aruma.wge(Post_Vaccine$new_deaths_per_million,d=1,s=7, phi = arima_s7_d1$phi,n.ahead = 180)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,276), main = "Next 180 Day forecast with p=7, q=0, s=7 and d=1")
lines(seq(97,276,1),for_aruma2_s7d1_f180$f, col = "red")
for_aruma2_s7d1_f380 = fore.aruma.wge(Post_Vaccine$new_deaths_per_million,d=1,s=7,  phi = arima_s7_d1$phi,n.ahead = 380)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,476), main = "Next 380 Day forecast with p=7, q=0, s=7 and d=1")
lines(seq(97,476,1),for_aruma2_s7d1_f380$f, col = "red")
```

### Multivariate - model with only vaccine variable ####

```{r, message=F, warning=F}
# Considering Only new_vaccinations_smoothed_per_million
PVsmall = Post_Vaccine[1:80,]
mv_fit0 <- lm( new_deaths_per_million~ new_vaccinations_smoothed_per_million  ,data = PVsmall)
aic.wge(mv_fit0$residuals, p=0:8,q=0) #AIC picks 8,0 ; aic =0.8019881
fit0 = arima(PVsmall$new_deaths_per_million, order=c(8,0,0), xreg = PVsmall[,c(7)] )
fit0 #AIC 295.99; Only new_vaccinations_smoothed_per_million looks significant
acf(fit0$residuals) #appear to be white noise
#ljung test for white noise of residuals
ltest = ljung.wge(fit0$residuals) #null hypothesis = white noise, alternate- not white noise. pval = 0.975. Here we FTR null hypothesis, so this is white noise
ltest
preds_mv0 = predict(fit0, newxreg = (Post_Vaccine$new_vaccinations_smoothed_per_million[81:96]   ))
ASE_mv0 = mean((Post_Vaccine$new_deaths_per_million[81:96] - preds_mv0$pred)^2)
ASE_mv0   #1.896324
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,96), ylab = "new_deaths_per_million", main = "Multivariate -Last 10 day forecast with Only new_vaccinations")
lines(seq(81,96,1),preds_mv0$pred, col = "red") # Even thought the ASE is low, plot doesn't capture the trend very well.
phi = aic.wge(mv_fit0$residuals)
m1_resids_0 = fore.arma.wge(mv_fit0$residuals, phi = phi$phi, n.ahead = 14)
#Predicting future forecasts - 14
Pred_nc_10_m1 = fore.aruma.wge(Post_Vaccine$new_cases_per_million, n.ahead = 14)
Pred_nv_10_m1 = fore.aruma.wge(Post_Vaccine$new_vaccinations_smoothed_per_million, n.ahead = 14)
Pred_hs_10_m1 = fore.aruma.wge(Post_Vaccine$hosp_patients_per_million, n.ahead = 14)
Pred_ic_10_m1 = fore.aruma.wge(Post_Vaccine$icu_patients_per_million, n.ahead = 14)
next10_m1 = data.frame(new_cases_per_million=Pred_nc_10_m1$f, new_vaccinations_smoothed_per_million=Pred_nv_10_m1$f,hosp_patients_per_million=Pred_hs_10_m1$f,icu_patients_per_million=Pred_ic_10_m1$f)
#get predictions
preds_m1_10 = predict(mv_fit0, newdata = next10_m1)
Preds_m1_10_final = preds_m1_10 + m1_resids_0$f
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,110), ylab = "new_deaths_per_million", main = "Multivariate - Next 14 day forecast with Only new_vaccinations")
lines(seq(97,110,1),Preds_m1_10_final, col = "red")
#Predicting future forecasts - 40
m1_resids_40 = fore.arma.wge(mv_fit0$residuals, phi = phi$phi, n.ahead = 40)
Pred_nc_40_m1 = fore.aruma.wge(Post_Vaccine$new_cases_per_million, n.ahead = 40)
Pred_nv_40_m1 = fore.aruma.wge(Post_Vaccine$new_vaccinations_smoothed_per_million, n.ahead = 40)
Pred_hs_40_m1 = fore.aruma.wge(Post_Vaccine$hosp_patients_per_million, n.ahead = 40)
Pred_ic_40_m1 = fore.aruma.wge(Post_Vaccine$icu_patients_per_million, n.ahead = 40)
next40_m1 = data.frame(new_cases_per_million=Pred_nc_40_m1$f, new_vaccinations_smoothed_per_million=Pred_nv_40_m1$f,hosp_patients_per_million=Pred_hs_40_m1$f,icu_patients_per_million=Pred_ic_40_m1$f)
#get predictions
preds_m1_40 = predict(mv_fit0, newdata = next40_m1)
Preds_m1_40_final = preds_m1_40 + m1_resids_0$f
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,136), ylab = "new_deaths_per_million", main = "Multivariate - Next 40 day forecast with Only new_vaccinations")
lines(seq(97,136,1),Preds_m1_40_final, col = "red")
```


### Multivariate - model with all explanatory variables ####


```{r, message=F, warning=F}
#Multivariate - Model 1 - No trend and no lags
#PVsmall = Post_Vaccine[1:80,]
mv_fit1 <- lm( new_deaths_per_million~ new_cases_per_million+ icu_patients_per_million+hosp_patients_per_million+new_vaccinations_smoothed_per_million  ,data = PVsmall)
aic.wge(mv_fit1$residuals, p=0:8,q=0) #AIC picks 7,0 ; aic =0.8019881
fit1 = arima(PVsmall$new_deaths_per_million, order=c(7,0,0), xreg = PVsmall[,c(3,5,6,7)] )
fit1 #AIC 281.45; Only new_cases_per_million looks significant
acf(fit1$residuals) #appear to be white noise
#ljung test for white noise of residuals
ltest = ljung.wge(fit1$residuals) #null hypothesis = white noise, alternate- not white noise. pval = 0.975. Here we FTR null hypothesis, so this is white noise
ltest
#Forecast model 1
preds_mv1 = predict(fit1, newxreg = cbind(Post_Vaccine$new_cases_per_million[81:96], Post_Vaccine$icu_patients_per_million[81:96],Post_Vaccine$hosp_patients_per_million[81:96],Post_Vaccine$new_vaccinations_smoothed_per_million[81:96]))
ASE_mv1 = mean((Post_Vaccine$new_deaths_per_million[81:96] - preds_mv1$pred)^2)
ASE_mv1 #2.59

#Multivariate - No trend and no lags with only new_cases_per_million
mv_fit1_2 <- lm( new_deaths_per_million~ new_cases_per_million  ,data = PVsmall)
aic.wge(mv_fit1_2$residuals, p=0:8,q=0) #AIC picks 7,0 ; aic =0.471746
fit1_2 = arima(PVsmall$new_deaths_per_million, order=c(7,0,0), xreg = PVsmall[,c(3)] )
fit1_2 #AIC 281.45; Only new_cases_per_million looks significant
acf(fit1_2$residuals) #appear to be white noise
#ljung test for white noise of residuals
ltest = ljung.wge(fit1_2$residuals) #null hypothesis = white noise, alternate- not white noise. pval = 0.975. Here we FTR null hypothesis, so this is white noise
ltest
#Forecast model 1_2
preds_mv1_2 = predict(fit1_2, newxreg = cbind(Post_Vaccine$new_cases_per_million[81:96]  ))
ASE_mv1_2 = mean((Post_Vaccine$new_deaths_per_million[81:96] - preds_mv1_2$pred)^2)
ASE_mv1_2 #3.736719

#Model 1 - No trend and no lags has the best ASE so far.


#Multivariate - Model 2 - Trend but  no lags
t = seq(1:96)
mv_fit2 <- lm( new_deaths_per_million~ t[1:80]+new_cases_per_million+ icu_patients_per_million+hosp_patients_per_million+new_vaccinations_smoothed_per_million  ,data = PVsmall)
aic.wge(mv_fit2$residuals, p=0:8,q=0) #AIC picks 7,0, aic = 0.6252096
fit2 = arima(PVsmall$new_deaths_per_million, order=c(7,0,0), xreg = cbind(t[1:80],PVsmall[,c(3,5,6,7)]) )
fit2 #AIC aic = 274.31; time, hosp_patients_per_million looks insignificant,
acf(fit2$residuals) #appear to be white noise
#ljung test for white noise of residuals
ltest2 = ljung.wge(fit2$residuals) #null hypothesis = white noise, alternate- not white noise. pval = 0.975. Here we FTR null hypothesis, so this is white noise
ltest2
#Forecast model 2
preds_mv2 = predict(fit2, newxreg = cbind(t[81:96], Post_Vaccine$new_cases_per_million[81:96], Post_Vaccine$icu_patients_per_million[81:96],Post_Vaccine$hosp_patients_per_million[81:96],Post_Vaccine$new_vaccinations_smoothed_per_million[81:96]))
ASE_mv2 = mean((Post_Vaccine$new_deaths_per_million[81:96] - preds_mv2$pred)^2)
ASE_mv2 #5.407704

#Take hosp_patients_per_million out
mv_fit2_1 <- lm( new_deaths_per_million~ t[1:80]+new_cases_per_million+ icu_patients_per_million +new_vaccinations_smoothed_per_million  ,data = PVsmall)
aic.wge(mv_fit2_1$residuals, p=0:8,q=0) #AIC picks 7,0, aic = 0.6189045
fit2_1 = arima(PVsmall$new_deaths_per_million, order=c(7,0,0), xreg = cbind(t[1:80],PVsmall[,c(3,5,7)]) )
fit2_1 #AIC aic = 274.31; time, new_vaccinations_smoothed_per_million looks insignificant
#Forecast model 2_1
preds_mv2_1 = predict(fit2_1, newxreg = cbind(t[81:96], Post_Vaccine$new_cases_per_million[81:96], Post_Vaccine$icu_patients_per_million[81:96],Post_Vaccine$new_vaccinations_smoothed_per_million[81:96]  ))
ASE_mv2_1 = mean((Post_Vaccine$new_deaths_per_million[81:96] - preds_mv2_1$pred)^2)
ASE_mv2_1  #6.364181

#Take new_vaccinations_smoothed_per_million out
mv_fit2_2 <- lm( new_deaths_per_million~ t[1:80]+new_cases_per_million+ icu_patients_per_million   ,data = PVsmall)
aic.wge(mv_fit2_2$residuals, p=0:8,q=0) #AIC picks 7,0, aic = 0.5544063
fit2_2 = arima(PVsmall$new_deaths_per_million, order=c(7,0,0), xreg = cbind(t[1:80],PVsmall[,c(3,5)]))
fit2_2 #AIC aic = 275.97; time, AIC increased a little bit
acf(fit2_2$residuals) #appear to be white noise
#ljung test for white noise of residuals
ltest2_2 = ljung.wge(fit2_2$residuals) #null hypothesis = white noise, alternate- not white noise. pval = 0.975. Here we FTR null hypothesis, so this is white noise
ltest2_2

#Forecast model 2_1
preds_mv2_2 = predict(fit2_2, newxreg = cbind(t[81:96], Post_Vaccine$new_cases_per_million[81:96], Post_Vaccine$icu_patients_per_million[81:96]))
ASE_mv2_2 = mean((Post_Vaccine$new_deaths_per_million[81:96] - preds_mv2_2$pred)^2)
ASE_mv2_2  #6.53061 No improvememnt
#Model 1 - No trend and no lags has the best ASE so far.
#Multivariate - Model 2 - Trend and lags
#Lagged variables - all
ccf(Post_Vaccine$new_cases_per_million,Post_Vaccine$new_deaths_per_million) #7
ccf(Post_Vaccine$icu_patients_per_million,Post_Vaccine$new_deaths_per_million) #1
ccf(Post_Vaccine$hosp_patients_per_million,Post_Vaccine$new_deaths_per_million) #1
ccf(Post_Vaccine$new_vaccinations_smoothed_per_million,Post_Vaccine$new_deaths_per_million) #8
Post_Vaccine$new_cases_per_million_7 <- dplyr::lag(Post_Vaccine$new_cases_per_million,7)
Post_Vaccine$icu_patients_per_million_1 <- dplyr::lag(Post_Vaccine$icu_patients_per_million,1)
Post_Vaccine$hosp_patients_per_million_1 <- dplyr::lag(Post_Vaccine$hosp_patients_per_million,1)
Post_Vaccine$new_vaccinations_smoothed_per_million_8 <- dplyr::lag(Post_Vaccine$new_vaccinations_smoothed_per_million,8)
PVsmall = Post_Vaccine[1:80,]
mv_fit3 <- lm( new_deaths_per_million~ t[1:80]+new_cases_per_million_7+ icu_patients_per_million_1+hosp_patients_per_million_1+new_vaccinations_smoothed_per_million_8  ,data = PVsmall)
aic.wge(mv_fit3$residuals, p=0:8,q=0) #AIC picks 8,0, aic = 0.8735
fit3 = arima(PVsmall$new_deaths_per_million, order=c(8,0,0), xreg = cbind(PVsmall[,c(8,9,10,11)]) )
fit3 #AIC aic = 263.57; time, icu_patients_per_million_1 and hosp_patients_per_million_1 looks sig,
acf(fit3$residuals[9:80]) # does not appear to be white noise
ltest3 = ljung.wge(fit3$residuals)
ltest3 #FTR.There is not enough evidence to suggest that the residuals are serailly correlated. null hypothesis = white noise, alternate- not white noise. pval = 0.975. Here we FTR null hypothesis, so this is white noise

#Lagged variables - icu_patients_per_million_1 and hosp_patients_per_million_1 - Best aic so far with white noise residuals.
mv_fit3_1 <- lm( new_deaths_per_million~ t[1:80]+new_cases_per_million+ icu_patients_per_million_1+hosp_patients_per_million_1+new_vaccinations_smoothed_per_million  ,data = PVsmall)
aic.wge(mv_fit3_1$residuals, p=0:8,q=0) #AIC picks 7,0, aic = 0.5299
fit3_1 = arima(PVsmall$new_deaths_per_million, order=c(7,0,0), xreg = cbind(t[1:80],PVsmall[,c(3,9,10,7)]) )
fit3_1 #AIC= time, icu_patients_per_million_1 and hosp_patients_per_million_1 looks sig,
acf(fit3_1$residuals[8:80]) # appears to be white noise
ltest3_1 = ljung.wge(fit3_1$residuals) #FTR. There is not enough evidence to suggest that the residuals are serailly correlated. null hypothesis = white noise, alternate- not white noise. pval = 0.5670672 Here we FTR null hypothesis, so this is white noise
#Forecast model 2_1
preds_mv3_1 = predict(fit3_1, newxreg = cbind(t[81:96], Post_Vaccine$new_cases_per_million[81:96], Post_Vaccine$icu_patients_per_million_1[81:96],  Post_Vaccine$hosp_patients_per_million_1[81:96] , Post_Vaccine$new_vaccinations_smoothed_per_million[81:96]))
ASE_mv3_1 = mean((Post_Vaccine$new_deaths_per_million[81:96] - preds_mv3_1$pred)^2)
ASE_mv3_1  #4.948167
#As Model 1 (no trend, no lag) has the best ASE. We will forecast future values using this model.
m1_fit = lm(new_deaths_per_million~new_cases_per_million+new_vaccinations_smoothed_per_million+hosp_patients_per_million+icu_patients_per_million, data = Post_Vaccine)
phi = aic.wge(m1_fit$residuals)
m1_resids_14 = fore.arma.wge(m1_fit$residuals, phi = phi$phi, n.ahead = 14)

#Predicting future forecasts - 14
Pred_nc_10_m1 = fore.aruma.wge(Post_Vaccine$new_cases_per_million, n.ahead = 14)
Pred_nv_10_m1 = fore.aruma.wge(Post_Vaccine$new_vaccinations_smoothed_per_million, n.ahead = 14)
Pred_hs_10_m1 = fore.aruma.wge(Post_Vaccine$hosp_patients_per_million, n.ahead = 14)
Pred_ic_10_m1 = fore.aruma.wge(Post_Vaccine$icu_patients_per_million, n.ahead = 14)
next10_m1 = data.frame(new_cases_per_million=Pred_nc_10_m1$f, new_vaccinations_smoothed_per_million=Pred_nv_10_m1$f,hosp_patients_per_million=Pred_hs_10_m1$f,icu_patients_per_million=Pred_ic_10_m1$f)

#get predictions
preds_m1_10 = predict(m1_fit, newdata = next10_m1)
Preds_m1_10_final = preds_m1_10 + m1_resids_14$f
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,110), ylab = "new_deaths_per_million", main = "Next 14 day new_deaths_per_million Forecast")
lines(seq(97,110,1),Preds_m1_10_final, col = "red")

#Predicting future forecasts - 40
m1_resids_40 = fore.arma.wge(m1_fit$residuals, phi = phi$phi, n.ahead = 40)
Pred_nc_40_m1 = fore.aruma.wge(Post_Vaccine$new_cases_per_million, n.ahead = 40)
Pred_nv_40_m1 = fore.aruma.wge(Post_Vaccine$new_vaccinations_smoothed_per_million, n.ahead = 40)
Pred_hs_40_m1 = fore.aruma.wge(Post_Vaccine$hosp_patients_per_million, n.ahead = 40)
Pred_ic_40_m1 = fore.aruma.wge(Post_Vaccine$icu_patients_per_million, n.ahead = 40)
next40_m1 = data.frame(new_cases_per_million=Pred_nc_40_m1$f, new_vaccinations_smoothed_per_million=Pred_nv_40_m1$f,hosp_patients_per_million=Pred_hs_40_m1$f,icu_patients_per_million=Pred_ic_40_m1$f)

#get predictions
preds_m1_40 = predict(m1_fit, newdata = next40_m1)
Preds_m1_40_final = preds_m1_40 + m1_resids_40$f
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,136), ylab = "new_deaths_per_million", main = "Next 40 day new_deaths_per_million Forecast")
lines(seq(97,136,1),Preds_m1_40_final, col = "red")
```


### VAR MODEL 1 with no lagged variables ####



```{r, message=F, warning=F}
PVsmall = Post_Vaccine[1:80,]
VAR_PV = VAR(cbind(PVsmall$new_deaths_per_million,PVsmall$new_cases_per_million,PVsmall$icu_patients_per_million,PVsmall$hosp_patients_per_million,PVsmall$new_vaccinations_smoothed_per_million),lag.max = 8, type = "both")
pred = predict(VAR_PV,n.ahead = 16)
plot(Post_Vaccine$new_deaths_per_million, type = "l")
lines(seq(81,96,1),pred$fcst$y1[,1],col = "red")
ASE = mean((Post_Vaccine$new_deaths_per_million[81:96] - pred$fcst$y1[1:16])^2)
ASE 
pred_14 = predict(VAR_PV,n.ahead = 30)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,110),ylab = "new_deaths_per_million", main = "VAR MODEL 1 - Next 14 day new_deaths_per_million Forecast")
lines(seq(97,110),tail(pred_14$fcst$y1[,1],14), col = "red")
pred_40 = predict(VAR_PV,n.ahead = 56)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,136), ylab = "new_deaths_per_million", main = "VAR MODEL 1 - Next 40 day new_deaths_per_million Forecast")
lines(seq(97,136,1),tail(pred_40$fcst$y1[,1],40), col = "red")
```


### VAR MODEL 2 - with lagged variables (Hospital and ICU) ###


```{r, message=F, warning=F}
#####lagged icu_1, hosp_1######
VAR_PV = VAR(cbind(PVsmall$new_deaths_per_million[2:80],PVsmall$new_cases_per_million[2:80],PVsmall$icu_patients_per_million_1[2:80],PVsmall$hosp_patients_per_million_1[2:80],PVsmall$new_vaccinations_smoothed_per_million[2:80]),lag.max = 8, type = "both")
pred = predict(VAR_PV,n.ahead = 16)
plot(Post_Vaccine$new_deaths_per_million, type = "l")
lines(seq(81,96,1),pred$fcst$y1[,1],col = "red")
ASE = mean((Post_Vaccine$new_deaths_per_million[81:96] - pred$fcst$y1[1:16])^2)
ASE 
pred_14 = predict(VAR_PV,n.ahead = 30)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,110),ylab = "new_deaths_per_million", main = "VAR MODEL 2 - Next 14 day new_deaths_per_million Forecast")
lines(seq(97,110,1),tail(pred_14$fcst$y1[,1],14), col = "red")
pred_40 = predict(VAR_PV,n.ahead = 56)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,136),ylab = "new_deaths_per_million", main = "VAR MODEL 2 - Next 14 day new_deaths_per_million Forecast")
lines(seq(97,136,1),tail(pred_40$fcst$y1[,1],40), col = "red")
#Model entire data for future forecasting
VAR_PV_full = VAR(cbind(Post_Vaccine$new_deaths_per_million[2:96],Post_Vaccine$new_cases_per_million[2:96],Post_Vaccine$icu_patients_per_million_1[2:96],Post_Vaccine$hosp_patients_per_million_1[2:96],Post_Vaccine$new_vaccinations_smoothed_per_million[2:96]),lag.max = 8, type = "both")
pred_14 = predict(VAR_PV_full,n.ahead = 30)
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,110),ylab = "new_deaths_per_million", main = "VAR MODEL 2 - Next 14 day new_deaths_per_million Forecast")
lines(seq(97,110,1),tail(pred_14$fcst$y1[,1],14), col = "red")
pred_40 = predict(VAR_PV_full,n.ahead = 56) 
plot(as.numeric(Post_Vaccine$new_deaths_per_million), type = "l", xlim = c(1,136),ylab = "new_deaths_per_million", main = "VAR MODEL 2 - Next 40 day new_deaths_per_million Forecast")
lines(seq(97,136,1),tail(pred_40$fcst$y1[,1],40), col = "red")
## We will proceed with Model 2 - with lagged variables ##
```


### MLP With Vaccine Variable Only####


```{r, message=F, warning=F}
#Build the model
Vacc_Train = Post_Vaccine[1:80,]
Vacc_Test = Post_Vaccine[81:96,]
Vacc_xreg = data.frame(Vaccine = ts(Vacc_Train$new_vaccinations_smoothed_per_million))
set.seed(2)
Vacc_fit.mlp1 = mlp(ts(Vacc_Train$new_deaths_per_million) ,reps = 50,comb = "median", xreg = Vacc_xreg ) 
Vacc_fit.mlp1 #mse 1.168
plot(Vacc_fit.mlp1)
fore_vacc_df_exp = data.frame(Vaccine = ts(Post_Vaccine$new_vaccinations_smoothed_per_million))
Vacc_fore.mlp1 = forecast(Vacc_fit.mlp1, h = 16, xreg = fore_vacc_df_exp) 
MLP1_ASE = mean((Vacc_Test$new_deaths_per_million- Vacc_fore.mlp1$mean)^2)
MLP1_ASE 
plot(Vacc_fore.mlp1)
#New Vacc - 14 day forecast on this explanatory variable
fit.mlp.NV = mlp(ts(Post_Vaccine$new_vaccinations_smoothed_per_million),reps = 50,comb = "median")
fore.mlp2.14.NV = forecast(fit.mlp.NV, h = 14)
plot(fore.mlp2.14.NV) 

PV_DF_fore = data.frame(new_vaccinations_smoothed_per_million =ts(c(Post_Vaccine$new_vaccinations_smoothed_per_million,fore.mlp2.14.NV$mean)))
fit.mlp = mlp(ts(Post_Vaccine$new_deaths_per_million),reps = 50,comb = "median",xreg = PV_DF_fore)
fore.mlp = forecast(fit.mlp, h = 14, xreg = PV_DF_fore)
plot(fore.mlp)

#New Vacc - 40 day forecast on this explanatory variable
fit.mlp.NV = mlp(ts(Post_Vaccine$new_vaccinations_smoothed_per_million),reps = 50,comb = "median")
fore.mlp2.40.NV = forecast(fit.mlp.NV, h = 40)
plot(fore.mlp2.40.NV) 
PV_DF_fore = data.frame(new_vaccinations_smoothed_per_million =ts(c(Post_Vaccine$new_vaccinations_smoothed_per_million,fore.mlp2.40.NV$mean)))
fore.vacc_40 = forecast(fit.mlp, h = 40, xreg = PV_DF_fore)
plot(fore.vacc_40)
```


### MLP With all explanatory variables####

```{r, message=F, warning=F}
VC_train = Post_Vaccine[1:80,]
VC_test = Post_Vaccine[81:96,]
all_exp_xreg = data.frame(Vaccine = ts(t = ts(seq(1:80)),VC_train$new_vaccinations_smoothed_per_million), NewCase = ts(VC_train$new_cases_per_million), Hosp = ts(VC_train$hosp_patients_per_million), ICU = ts(VC_train$icu_patients_per_million))
set.seed(2)
allexp_fit_mlp2 = mlp(ts(VC_train$new_deaths_per_million), reps = 50, comb = "median",  xreg = all_exp_xreg )
plot(allexp_fit_mlp2)
fore_all_exp_df = data.frame(Vaccine = ts(Post_Vaccine$new_vaccinations_smoothed_per_million), NewCase = ts(Post_Vaccine$new_cases_per_million), Hosp = ts(Post_Vaccine$hosp_patients_per_million), ICU = ts(Post_Vaccine$icu_patients_per_million))
allexp_fore.m2 = forecast(allexp_fit_mlp2, h = 16, xreg = fore_all_exp_df) 
MLP2_ASE = mean((VC_test$new_deaths_per_million- allexp_fore.m2$mean)^2)
MLP2_ASE #1.43
#Forecast - fitting on last 16 days
plot(seq(1,96,1), Post_Vaccine$new_deaths_per_million, type = "l",xlim = c(0,96), ylab = "Death Rate Per Million", main = "forecast plot for last 6 days")
lines(seq(81,96), allexp_fore.m2$mean, type = "l", col = "red")
### Forecast on future short term (14 days) and long term (40 days) ###
#New Case Forecast
fit.mlp.NC = mlp(ts(Post_Vaccine$new_cases_per_million),reps = 50,comb = "median")
fore.mlp2.14.NC = forecast(fit.mlp.NC, h = 14)
fore.mlp2.40.NC = forecast(fit.mlp.NC, h = 40)
plot(fore.mlp2.14.NC) #for short term
plot(fore.mlp2.40.NC) #for long term
#New Vacc
fit.mlp.NV = mlp(ts(Post_Vaccine$new_vaccinations_smoothed_per_million),reps = 50,comb = "median")
fore.mlp2.14.NV = forecast(fit.mlp.NV, h = 14)
fore.mlp2.40.NV = forecast(fit.mlp.NV, h = 40)
plot(fore.mlp2.14.NV) #for short term
plot(fore.mlp2.40.NV) #for long term
#New Hos
fit.mlp.HS = mlp(ts(Post_Vaccine$hosp_patients_per_million),reps = 50,comb = "median")
fore.mlp2.14.HS = forecast(fit.mlp.HS, h = 14)
fore.mlp2.40.HS = forecast(fit.mlp.HS, h = 40)
plot(fore.mlp2.14.HS) #for short term
plot(fore.mlp2.40.HS) #for long term
#New ICU
fit.mlp.IC = mlp(ts(Post_Vaccine$icu_patients_per_million),reps = 50,comb = "median")
fore.mlp2.14.IC = forecast(fit.mlp.HS, h = 14)
fore.mlp2.40.IC = forecast(fit.mlp.HS, h = 40)
plot(fore.mlp2.14.IC) #for short term
plot(fore.mlp2.40.IC) #for long term
#Forecast 14 days
PV_DF_fore1 = data.frame(t = ts(seq(1:110)),new_cases_per_million = ts(c(Post_Vaccine$new_cases_per_million,fore.mlp2.14.NC$mean)), icu_patients_per_million = ts(c(Post_Vaccine$icu_patients_per_million,fore.mlp2.14.IC$mean)),
                        hosp_patients_per_million = ts(c(Post_Vaccine$hosp_patients_per_million,fore.mlp2.14.HS$mean)), new_vaccinations_smoothed_per_million =ts(c(Post_Vaccine$new_vaccinations_smoothed_per_million,fore.mlp2.14.NV$mean)))
fit.mlp2_14 = mlp(ts(Post_Vaccine$new_deaths_per_million),reps = 50,comb = "median",xreg = PV_DF_fore1)
fore.mlp2_14 = forecast(fit.mlp2_14, h = 14, xreg = PV_DF_fore1)
plot(fore.mlp2_14)
#Forecast 40 days
PV_DF_fore = data.frame(t = ts(seq(1:136)),new_cases_per_million = ts(c(Post_Vaccine$new_cases_per_million,fore.mlp2.40.NC$mean)), icu_patients_per_million = ts(c(Post_Vaccine$icu_patients_per_million,fore.mlp2.40.IC$mean)),
                        hosp_patients_per_million = ts(c(Post_Vaccine$hosp_patients_per_million,fore.mlp2.40.HS$mean)), new_vaccinations_smoothed_per_million =ts(c(Post_Vaccine$new_vaccinations_smoothed_per_million,fore.mlp2.40.NV$mean)))
fit.mlp2_40 = mlp(ts(Post_Vaccine$new_deaths_per_million),reps = 50,comb = "median",xreg = PV_DF_fore)
fore.mlp2_40 = forecast(fit.mlp2_40, h = 40, xreg = PV_DF_fore)
plot(fore.mlp2_40)
```



### Ensemble Model with NN (all variables) Model and VAR 2####

```{r, message=F, warning=F}

#preds_mv0$pred : multivariate with only vaccine
#allexp_fore.m2$mean: mlp with all variables
ensemble  = (preds_mv0$pred + allexp_fore.m2$mean)/2

#Plot
plot(seq(1,96,1), Post_Vaccine$new_deaths_per_million, type = "l",xlim = c(0,96), ylab = "Death Rate Per Million", main = "Ensemble Model with NN and Multivariate models")
lines(seq(81,96,1), ensemble, type = "l", col = "green")

ASE_en = mean((Post_Vaccine$new_deaths_per_million[81:96] - ensemble)^2)
ASE_en #1.590614

#for_aruma2_s7d1_16: Univariate 16 last n
#allexp_fore.m2$mean: mlp with all variables

#Univariate Model 3 and MLP with all variables
ensemble2  = (for_aruma2_s7d1_16$f + allexp_fore.m2$mean)/2

#Plot
plot(seq(1,96,1), Post_Vaccine$new_deaths_per_million, type = "l",xlim = c(0,96), ylab = "Death Rate Per Million", main = "Ensemble Model with NN (All Var) and ARIMA models")
lines(seq(81,96,1), ensemble2, type = "l", col = "green")

ASE_en2 = mean((Post_Vaccine$new_deaths_per_million[81:96] - ensemble2)^2)
ASE_en2 #1.836393

```

